---
title: "Distributions homework"
output: html_notebook
---
## MVP

### Question 1

__Load the data, `clean_names()` and explore it.__

First load the required libraries

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
```
Read in the data and have a quick look.
```{r}
phone_data <- read_csv("data/20190928-items.csv") %>% 
  clean_names()

glimpse(phone_data)
```

The data has 9 columns and 792 rows. All columns appear to be classed correctly. Prices has some missing data.

Check missing data across all columns.

```{r}
phone_data %>% 
  summarise(across(everything(), ~sum(is.na(.x))))
```

The only column missing data is the `prices` column. This is missing 215 variables.
This column is also currently coded as a character column due to the $.
Reviewing the questions it appears this column is not required at this stage. Leave this column as it is for now. <br>

### Question 2

__Find the brand with the highest numbers of phones in the dataset.__

```{r}
phone_data %>% 
  group_by(brand) %>% 
  summarise(total = n()) %>% 
  slice_max(total, with_ties = TRUE)
```

Samsung has the most number of phones in the data set with 397 phones in total. This is just over half the phones in the `phone_data` data set. <br>

### Question 3

__For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution. Do you think the normal distribution provides a good model of these ratings?<br><br>[**Hint** - you will need to calculate the `mean()` and `sd()` of the ratings to plot the appropriate normal. Alternatively, investigate using the `fitdistr()` function in the `MASS` package]__

```{r}
#filter data for top brand
samsung_phones <- phone_data %>% 
  filter(brand == "Samsung")

#generate summary stats
samsung_phone_stats <- samsung_phones %>% 
  summarise(
    num = n(),
    mean = mean(rating),
    sd = sd(rating)
  )

#plot distribution
samsung_phones %>% 
  ggplot(aes(x = rating)) +
  geom_histogram(aes(y = ..density..), col = "white") +
  stat_function( #use this to plot the norm distrib over the histogram
    fun = dnorm,
    args = list(
      mean = samsung_phone_stats$mean,
      sd = samsung_phone_stats$sd
    ),
    col = "blue"
  )
```

The probability density of ratings for the top brand (Samsung) shows a slight left skew. The fitted normal distribution does not show a good fit to the data. There are some outliers seen in the plot as peaks which are found outside of the fitted normal distribution. <br>

## Question 4

We can be more quantitative in comparing the distribution of top brand ratings with a normal distribution. You calculated the `mean()` and `sd()` of your distribution in an earlier question. Use these values to calculate the *proportions* of ratings within one-, two- and three standard deviations of the mean. Compare these proportions with those you expect for a normal distribution. 

```{r}
#function to calculate proportion of data within x sd from the mean
prop_within_n_sd_calculator <- function(data, n){
  data %>% 
    filter(rating >= samsung_phone_stats$mean - n*samsung_phone_stats$sd,
         rating <= samsung_phone_stats$mean + n*samsung_phone_stats$sd) %>%
  summarise(prop_within_sd = n()/samsung_phone_stats$num) %>% 
    pull()
}

#generate table with expected proportions for a normal distribution
prop_table <- tibble(
within_1_sd = (pnorm(q = 1) - pnorm(q = -1)),
within_2_sd = (pnorm(q = 2) - pnorm(q = -2)),
within_3_sd = (pnorm(q = 3) - pnorm(q = -3))
)

#calc proportions from mean

prop_table_data <- tibble(
within_1_sd = prop_within_n_sd_calculator(samsung_phones, 1),
within_2_sd = prop_within_n_sd_calculator(samsung_phones, 2),
within_3_sd = prop_within_n_sd_calculator(samsung_phones, 3)
)

prop_table
prop_table_data
```

This data has a proportion of 0.79 of the data within 1 standard deviations away from the mean, a proportion of 0.92 of the data within 2 standard deviations from the mean and 0.98 of the data within 3 standard deviations of the mean. For an ideally normally distributed dataset with a mean of 0, a proportion of 0.68 of the data should be distributed within 1 standard deviation of the mean, a proportion of 0.95 of the data should be distributed within 2 standard deviations of the mean and a proportion of 0.99 of the data should be distributed within 3 standard deviations of the mean.

Taking all this information together, I would say this suggests that the distribution of the data is not normally distributed.