---
title: "ggplot Practice Questions - Homework"
output: html_notebook
---


```{r}
library(tidyverse)
loans <- read_csv("data/loans.csv")

head(loans)
```

# Question 1

```{r}
loans %>% 
  ggplot() +
  aes(x = loan_amnt, y = int_rate, colour = grade) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ grade)
```

These scatter plots show the loan amount on the x axis and the interest rate on the y axis. Each plot shows a separate loan grade. These plots illustrate that interest rate increases as the loan grade increases from A to G.

# Question 2

```{r}
loans %>% 
  group_by(grade) %>% 
  summarise(
    mean_loan = mean(loan_amnt),
    mean_int_rate = mean(int_rate)
  ) %>% 
  ggplot() +
  aes(x = grade, y = mean_loan, fill = mean_int_rate) +
  geom_col()
```

# Question 3
summarises how purpose, loan about, interest rate and grade are related. Is there one purpose that is significantly different from the others? How can you show this in your graph?

```{r}
loans %>% 
  group_by(grade, purpose) %>% 
  summarise(
    mean_loan = mean(loan_amnt),
    mean_int_rate = mean(int_rate)
  ) %>% 
  ggplot() +
  aes(x = grade, y = mean_loan, fill = mean_int_rate) +
  geom_col() +
  facet_wrap(~purpose)
```

```{r}
loans %>% 
  group_by(grade, purpose) %>% 
  summarise(
    mean_loan = mean(loan_amnt),
    mean_int_rate = mean(int_rate)
  ) %>% 
  ggplot() +
  aes(x = mean_loan, y = mean_int_rate, colour = purpose) +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = c("Other" = "#e7298a", "Credit Card" = "#b3e2cd", "Debt Consolidation" = "#fdcdac", "Home Improvement" = "#cbd5e8", "Small Business" = "#e6f5c9")) +
  theme_minimal() +
  labs(
    title = "Mean interest rate and mean loan rate",
    subtitle = "Other loan types",
    x = "Mean loan rate",
    y = "Mean interest rate"
  )
```

