---
title: "Intro to ggplot2 homework assignment"
author: "Anne Braae"
date: "20/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Question 1.**  

Load in the libraries you need, and look at the `backpack` data. Take note of variable names and dimensions. 

Load libraries
```{r}
library(tidyverse)
library(CodeClanData)
```
Load data and get dimensions
```{r}
backpack <- backpack
head(backpack)
summary(backpack)
dim(backpack)
```
  The backpack data set contains 100 observations (rows) and 9 variables (columns).
  The columns types appear to be classed correctly. Note that the major and sex columns are factor character columns.


**Question 2.**

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution. 

For this I would use a histogram.

```{r}

```



**Question 3.**  

Are there differences in backpack weight between sex?   


**Question 4.**  

What is the relationship between body weight and backpack weight?


**Question 5.**  

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.  


**Question 6.**  

What **proportion** of students report back pain?   


**Question 7.**  

Is this proportion different between men and women? 


**Question 8.**  

Now you've investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both `backpack_weight` and `body_weight` as well as the `ratio`.   

*Hint: faceting would be useful here*.   

# Extensions

<br>

**Question 1.** 

Does reported backpain differ across the year groups? Ignore year 0 in your data.  


**Question 2.** 

Is there a difference between how heavy the bags those in different years carry?



