---
title: "Clustering homework assignment"
output:
  html_document:
    df_print: paged
---

* You have been provided some customer data from a shopping centre. Do some exploratory analysis on the data. Comment on findings. 

```{r}
library(tidyverse)
library(factoextra)
library(corrplot)
library(broom)
```

```{r}
mall_customers <- read_csv("data/mall_customers.csv") %>% 
  janitor::clean_names()
```

```{r}
# check for missing data
any(is.na(mall_customers))

# count distinct data
mall_customers %>% 
  summarise(across(everything(), ~ n_distinct(.x)))
```
* set row_names to `customer_id`
* set gender to factor

```{r}
mall_customers <- mall_customers %>% 
  column_to_rownames("customer_id") %>%
  mutate(gender = as.factor(gender))
```

There are two genders which I have set as a factor character. The `customer_id` has been set as a row_name. There are no missing values. I will have a look at possible correlations in the numeric data after scaling the data.

```{r}
mall_customers_scale <- mall_customers %>% 
  mutate_if(is.numeric, scale)

mall_customers_scale_num <- mall_customers_scale %>% 
  select(is.numeric)

corrplot(cor(mall_customers_scale_num), method = "number", type = "lower")
```
Looks like in the numeric data age and spending_score_1_100 are moderately weakly correlated.

* We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers. 

```{r}
# make a single k-means cluster

cluster_mall <- kmeans(mall_customers_scale_num,
                       centers = 6,
                       nstart = 25)

cluster_summary <- tidy(cluster_mall,
     col.names = colnames(mall_customers_scale_num))
cluster_summary

glance(cluster_mall)
```
Looks like the data can be clustered. For 6 clusters the total within sum of squares is 133.199, and the number of iterations stops at 4. The within sum of squares for individual clusters is low (20) at 2 clusters and at 4 clusters.

plot income and spencing score!

```{r}
#ggplot(customers, aes(x = ann))
```


* Perform k-means clustering and chose a value of k.

```{r}
# set maximum numbers of clusters
k_max <- 20

k_clusters <- tibble(k = 1:k_max) %>%
  mutate(
    kclust = map(k, ~ kmeans(
      mall_customers_scale_num, .x, nstart = 25)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, mall_customers_scale)
  )

unnested_k_clusters <- k_clusters %>% 
  unnest(glanced)

unnested_k_clusters
```
```{r}
# look for the "elbow"
ggplot(unnested_k_clusters) +
  aes(x = k, y = tot.withinss) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(1,20, by = 1))
```

Looks like 2 or 4 clusters produces an elbow. Let's check another metric to select clusters.

```{r}
library(factoextra)
fviz_nbclust(mall_customers_scale_num,
             kmeans,
             method = "silhouette",
             nstart = 25)
```

The automatic output has selected 6 clusters, but I think you could argue for 2 given the line graph shown.

Let's have one last look at the final metric for selecting clusters, the gap statistic.

```{r}
fviz_nbclust(mall_customers_scale_num,
             kmeans,
             method = "gap_stat",
             nstart = 25)
```
Ok the gap statistic is also selecting 6 clusters. So we will go with 6 clusters as our chosen k value.

* Visualise the clustering for your chosen value of k.

```{r}
unnested_k_clusters %>% 
  unnest(cals = c(augmented)) %>% 
  filter(k == 6) %>% 
  ggplot() +
  aes(x = age, y = annual_income_k, colour = .cluster) +
  geom_point(aes(colour = .cluster))

unnested_k_clusters %>% 
  unnest(cals = c(augmented)) %>% 
  filter(k == 6) %>% 
  ggplot() +
  aes(x = age, y = spending_score_1_100, colour = .cluster) +
  geom_point(aes(colour = .cluster))

unnested_k_clusters %>% 
  unnest(cals = c(augmented)) %>% 
  filter(k == 6) %>% 
  ggplot() +
  aes(x = annual_income_k, y = spending_score_1_100, colour = .cluster) +
  geom_point(aes(colour = .cluster))
```

* Do you think the clustering seems a good fit for this data?

It seems ok. I would probably remove age from the clustering analysis and run the k-means clustering again.

* Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the 'Segmentation & clustering intro' lesson).

The clustering seems to work the best for annual income and age.
Spending and age do not really show good groupings. But spending and annual income do show good groupings.

I cannot find section 4.1 of the intro lesson? Maybe it has been relabeled?

However for the clustering for age and spending, there is a clear group, 3 with an above average spending score and a below average customer age who may be targeted with marketing as a useful group to increase spending at the mall. Looking at this cluster in income and spending, their income is also above average. I think this would be a good group of customers to target with incentives to come to the mall and would increase spending overall.

Labelling from Tom:

```{r}
# 5 appears to fit well
k5 <- clusterings %>% 
  unnest(cols = c(augmented)) %>% 
  filter(k == 5)

# get cluster info for centres
k5clust <- k5 %>% 
  select(kclust) %>% 
  pull()
```


```{r}
# notice how the centres are fit to the scaled data
k5clust[[1]]$centers
```


```{r}
# define function to remove scaling
remove_center_scaling <- function(centre_mat, scaled_df) {
  vars <- colnames(centre_mat)
  
  unscaled_centres <- sapply(
    vars,
    function(x) centre_mat[, x] * attr(scaled_df[[x]], 'scaled:scale') +
      attr(scaled_df[[x]], 'scaled:center'))
  
  return(as_tibble(unscaled_centres))
}
```

```{r}
# call function. remove scaling from centroids
unscaled_centres <- remove_center_scaling(
  k5clust[[1]]$centers, mall_customers_scale
  )
```

```{r}
ggplot(k5) + 
  geom_point(aes(x = annual_income_k, y = spending_score_1_100,
                 colour = .cluster)) +
    geom_text(data = unscaled_centres,
              aes(x = .data[["annual_income_k"]],
                  y = .data[["spending_score_1_100"]]),
              color = "black", size = 4, label = 1:5) +
  stat_ellipse(data = k5,
               aes(x = annual_income_k, y = spending_score_1_100,
                   colour = .cluster, fill = .cluster),
               geom = "polygon", alpha = 0.1)