---
title: "dplyr weekend homework"
output: html_notebook
---

#Load in libraries (update with additional libraries as needed)
```{r}
library(tidyverse)
library(janitor)
```


#Load in the data and initial investigation

```{r}
book <- read_csv("data/books.csv") 
book
glimpse(book)
names(book)

```
#Initial thoughts
Heading names seem fine apart from bookID. Variables appear to have been classed appropriately.

Questions I would like to work through:
How many different authors are there? Are there some publishing houses with more authors than others?
What is the language split between the authors/publishing houses?
  How many languages are there?
  Are there languages which are not english? Do I need to deal with these?
Link between page numbers and rating?
Anything going on between language and rating? (i.e. are eng and eng-us books rated higher and or read more?)

Also:
Cool error on load up!

Warning: 21 parsing failures.
 row            col           expected            actual             file
1570 title          delimiter or quote                   'data/books.csv'
1570 title          delimiter or quote I                 'data/books.csv'
3349 average_rating a double           Jr./Sam B. Warner 'data/books.csv'
3349 num_pages      a double           en-US             'data/books.csv'
3349 NA             12 columns         13 columns        'data/books.csv'
.... .............. .................. ................. ................
See problems(...) for more details.


#Investigating errors on load

```{r}
#Looking at error 1 on row 1570
slice(book, 1570)
```
This appears to be a strange title with extra quotation marks

```{r}
#Look at the next error row
slice(book, 3349)
```
Something strange is going on here. Looks like the data has been shifted 1 column to the right. Possible data entry issue.

I am going to try the error suggestion and adjust for quotes in the read in step. Let's see if I can remove some errors

```{r}
book <- read_csv("data/books.csv", quote = "") 
#after reading the parsing errors tried adding: quote = "" to the above
```


Warning: 12 parsing failures.

Fantastic! Now I only have 12 parsing failures instead of 21. 

I am going to try the error suggestion and run problems(book) to pull out the errors with the dataset into a new dataset, 'book_errors'. So I can easily look at the rows which are causing trouble.

```{r}
error_rows <- problems(book) %>%
  distinct(row) %>% 
  pull()
book_errors <- slice(book, error_rows)

```

These four rows contain data that has skipped a row, this is because of an extra comma in the authors column. I do not know how to deal with this by code. I will manually edit them and resave the csv as books_edit.csv to continue.

Reload the data and clean the names at the same time.

```{r}
book_edit <- read_csv("data/books_edit.csv", quote = "") %>%
  clean_names()

names(book_edit)
```
#Missing data
Now let's look at missing data in the remaining rows, summarised across the columns

```{r}
# writing this function to count na in a column
na_count <- function(col){
    return(sum(is.na(col)))
}

book_edit %>%
  summarise(across(everything(), na_count))
```

Hmm. There are no NAs! At least no NAs being recognised by R.

Ok so now I can recode some of the zeros I can reasonably assume to be NAs.

Specifically:
num_pages = 0 
average rating if average_rating = 0 and ratings_count > 0
ratings count if average_rating > 0 and ratings_count = 0


```{r}
book_edit_na <- book_edit %>% 
  mutate(num_pages = na_if(num_pages, 0),
         average_rating = if_else(average_rating == 0 & ratings_count > 0, NA_real_,
                                  as.numeric(average_rating)),
         ratings_count = if_else(ratings_count == 0 & average_rating > 0, NA_real_,
                    as.numeric(ratings_count))
         )

#count NAs introduced
book_edit_na %>%
  summarise(across(everything(), na_count))

#Resummarise cleaned dataset
dim(book_edit_na)
glimpse(book_edit_na)
head(book_edit_na)
```

#Data cleaning summary:
Books.csv has been loaded, parsing errors fixed, names cleaned, and missing values recoded. Cleaned dataset book_edit_na has 12 columns (variables) and 11127 rows (observations). Data has been recoded as missing and there are 76 missing page numbers (num_pages) and 55 missing ratings count (ratings_count). At this stage I will leave the NAs in the dataset, as they are only a small fraction of the total data in each column.

```{r}
book_edit_na %>% 
  distinct(authors)
```
 
Questions I would like to work through:
How many different authors are there? Are there some publishing houses with more authors than others?
What is the language split between the authors/publishing houses?
  How many languages are there?
  Are there languages which are not english? Do I need to deal with these?
Link between page numbers and rating?
Anything going on between language and rating? (i.e. are eng and eng-us books rated higher and or read more?)


#Looking at the authors
Questions I want to answer:
How many different authors are there? 
Who are the top ten authors (based on average rating)?
Who are the bottom ten authors (based on average rating)?
Are there some publishing houses with more authors than others?
```{r}
book_edit_na %>%
  distinct(authors) %>% 
  count()


```
There are 6642 different authors listed. HOWEVER I note some authors may be listed more than once if they are co authors. 
There are too many authors to examine individually. Let's look at the top ten based on average rating, and also page number.

```{r}
author_subset <- book_edit_na %>% 
  select(authors, title, ratings_count, num_pages)

reviewed_top_ten <- author_subset %>% 
  slice_max(ratings_count, n = 10)
reviewed_top_ten

longest_top_ten <- author_subset %>% 
  slice_max(num_pages, n = 10)
longest_top_ten

author_subset %>% 
  group_by(authors) %>% 
  summarise(mean_pgs = mean(num_pages)) %>% 
  slice_max(mean_pgs, n = 10)

author_subset %>%
  group_by(authors) %>% 
  mutate(total_books = distinct(title))
  summarise(total_books = distinct(title)) %>% 
  slice_max(total_books, n = 10)
  
test <- author_subset %>%
  group_by(authors)%>%
  distinct(authors)
test

#this counts the number of times an author has been mentioned.
author_subset %>%
  count(authors) %>%
```



book_edit_na %>%
  group_by(authors) %>% 
  select(authors, ratings_count, num_pages) %>% 
  summarise(across(.cols = everything(), sum)
    )
    
rated_top_ten <- book_edit_na %>%
  select(authors, title, average_rating, ratings_count, num_pages) %>% 
  slice_max(average_rating, n = 10)




#Looking at the languages
What is the language split between the authors/publishing houses?
  How many languages are there?
  Are there languages which are not english? Do I need to deal with these?

```{r}
book_edit_na %>% 
  distinct(language_code)
```
In the language_codes column english is coded three times eng (which is not a localisation language code), en-US, en-GB and en-CA.
For language_code, eng is not a localisation language code - recode to en.