---
title: "ggplot Homework"
output: html_notebook
---

Load libraries

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(janitor)
library(CodeClanData)
library(scales)
```

Read in required data and have a quick look

```{r}
qb_revenue_breakdown <- clean_names(qb_revenue_breakdown)
qb_revenue_breakdown

qb_monthly_sales <- clean_names(qb_monthly_sales)
qb_monthly_sales

qb_competitors <- clean_names(qb_competitors)
qb_competitors

qb_device_data <- clean_names(qb_device_data)
qb_device_data
```


# MVP

## Part 1

1. Take the data in the dataset `qb_revenue_breakdown` and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}
ggplot(qb_revenue_breakdown) +
  aes(x = product, y = revenue, fill = year) +
  geom_bar(stat = "identity")
```


2.	Make a line chart showing monthly sales of the "flagship product" - the 'KwikBit Unit 1' - over the last year. This data is inside `qb_monthly_sales`, but you will need to use `filter()` or `subset()` (base `R)`.

```{r}
qb_monthly_sales %>% 
  filter(cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot() +
  aes(x = date, y = amount) +
  geom_line() +
  scale_x_date(breaks = "1 month", date_labels = "%b %y")
```

3.	Make a line chart showing monthly revenue and costs over the last year. This data is also in `qb_monthly_sales`.	

```{r}
qb_monthly_sales %>% 
  filter(cashflow == "Cost") %>% 
  ggplot() +
  aes(x = date, y = amount) +
  geom_line() +
  scale_x_date(breaks = "1 month", date_labels = "%b %y")
# note was suggested to used date_breaks instead of breaks - date_breaks seems to be depreciated?
```

4.	Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use `geom_area`). This data is in `qb_competitors`.

```{r}
ggplot(qb_competitors) +
  aes(x = year, y = revenue, group = company, fill = company) +
  geom_area()
```

5. Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
ggplot(qb_competitors) +
  aes(x = year, y = revenue, group = company, colour = company) +
  geom_line() +
  geom_point()
```

6. Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in `qb_device_data`.

```{r}
qb_device_data %>% 
  mutate(tot_mins = hours+(mins/60)) %>% 
ggplot() +
  aes(x = tot_mins, y = counts, colour = device) +
  geom_line()+
  facet_grid(id ~ device)

#I am not sure this graph is right. Looks very busy.
```


## Part 2 

Take the plots that you produced in part one and now polish them by:

1. Adding appropriate labels
2. Changing the scales and coordinates when appropriate.
3. Applying a unified theme, which is described below:
  * Graphs should have white backgrounds, and use colour sparingly.
  * There should be faint grid lines.
  * Font sizes should be ~12pt although titles should be slightly larger and axis labels can be slightly smaller. 
  * All plots should use colours from the following company colour scheme.

```{r, echo=TRUE}
col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")
```

```{r}
qb_theme <- function(){
  theme(
    #add border
    panel.border = element_rect(colour = "#50434F", fill = NA, linetype = 1),
    #white background
    panel.background = element_rect(fill = "white"),
    #faint gridlines
    panel.grid.major = element_line(colour = "#B6A7B5", linetype = 8, size = 0.25),
    panel.grid.minor = element_blank(),
    #modify font sizes in axis and titles
    plot.title = element_text(colour = "#50434F", size = 14),
    plot.subtitle = element_text(colour = "#50434F", size = 12),
    axis.text = element_text(colour = "#50434F", size = 10),
    axis.title = element_text(colour = "#50434F", size = 12),
    axis.ticks = element_line(colour = "#50434F"),
    #modify font sizes in legend
    legend.title = element_text(colour = "#50434F", size = 12),
    legend.text = element_text(colour = "#50434F", size = 10),
    legend.background = element_rect(fill = "white"),
    legend.key = element_rect(fill = "white")
  )
}
```


### Qu 1 - part 2
```{r}
ggplot(qb_revenue_breakdown) +
  aes(x = product, y = revenue, fill = year) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = col_scheme) +
  labs(title = "\nSources of Revenue for 2018 and 2019\n",
       x = "\nProducts", y = "Revenue (£)\n", fill = "Year") +
  qb_theme()
```

### Qu 2 - part 2
```{r}
qb_monthly_sales %>% 
  filter(cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot() +
  aes(x = date, y = amount) +
  geom_line(colour = "#E89FE9") +
  scale_x_date(breaks = "1 month", date_labels = "%b %y") +
  labs(title = "\nMonthly KwickBit Unit 1 Sales",
       subtitle = "Sales of KwickBit Unit 1 from 1 Aug 2018 to 1 Jul 2019\n",
       x = "\nDate", y = "Total Sales (£)\n") +
  qb_theme()
```

### Qu 3 - part 2

```{r}
qb_monthly_sales %>% 
  filter(cashflow != "Kwikbit Unit 1 Sales") %>% 
  ggplot() +
  aes(x = date, y = amount, group = cashflow, colour = cashflow) +
  geom_line() +
  scale_colour_manual(values = col_scheme[3:4]) +
  scale_x_date(breaks = "2 months", date_labels = "%b %y") +
  scale_y_continuous(breaks = c(1500, 2000, 2500, 3000), 
                     limits = c(1200, 3200)) +
  labs(title = "\nMonthly Revenue and Costs",
       subtitle = "Amounts per Month from 1 Aug 2018 to 1 Jul 2019\n",
       x = "\nDate", y = "Amount (£)\n", colour = "") +
  qb_theme()
```

### Qu 4 - part 2

```{r}
ggplot(qb_competitors) +
  aes(x = year, y = revenue, group = company, fill = company) +
  geom_area() +
  scale_fill_manual(values = col_scheme[2:5]) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "\nAnnual Sales of Personal Fitness Trackers",
       subtitle = "Sales per Company over the last 5 years\n",
       x = "\nYear", y = "Revenue (£)\n", fill = "Company") +
  qb_theme()
```

### Qu 5 - part 2

```{r}
ggplot(qb_competitors) +
  aes(x = year, y = revenue, group = company, colour = company) +
  geom_line() +
  geom_point() +
  scale_colour_manual(values = col_scheme[2:5]) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "\nAnnual Sales of Personal Fitness Trackers",
       subtitle = "Sales per Company over the last 5 years\n",
       x = "\nYear", y = "Revenue (£)\n", colour = "Company") +
  qb_theme()
```

### Qu 6 - part 2

```{r}
qb_device_data %>% 
  mutate(tot_mins = hours+(mins/60)) %>% 
ggplot() +
  aes(x = tot_mins, y = counts, colour = device) +
  geom_line()+
  facet_grid(id ~ device) +  
  scale_colour_manual(values = col_scheme[1:4]) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "\nSteps per Device over 24 hours",
       subtitle = "Steps measured per device for the same five people\n",
       x = "\nTime (in hours)", y = "Steps\n", colour = "Company") +
  qb_theme()
```


# Extension

Try out different ways of plotting these three datasets. Which graphs do you think are particularly effective at showing QikBit's success?
