---
title: "Probability Homework"
output: html_notebook
---

#1 MVP - The titanic

Load in the libraries
```{r, message=FALSE, warning=FALSE}
library(prob)
library(gtools)
library(tidyverse)
```

Load in the data

```{r}
titanic <- read_csv("data/titanic.csv")
```

Have a look at distinct class variables

```{r}
titanic %>% 
  distinct(class)
```
7 distinct classes, 3 for passengers (1st, 2nd and 3rd) and the rest for the crew.

Add a crew column to filter the data into passengers and crew.

```{r}
passengers <- c("3rd", "2nd", "1st")

titanic <- titanic %>% 
  mutate(crew = !class %in% passengers)
```

### Calculate the conditional probability that crew members were more likely to have survived than passengers. 

prob( survived ∣ crew )
```{r}
p_survived_given_crew <- titanic %>% 
  filter(crew == TRUE) %>% 
  summarise(p = sum(survived == "yes")/n())

p_survived_given_crew
```

The probability of surviving GIVEN being crew is 0.237.

Check:

Probability of survival AND being crew:
```{r}
p_survival_crew <- titanic %>% 
  filter(crew == TRUE,
         survived == "yes") %>% 
  summarise(p = n()/nrow(titanic))
p_survival_crew
```

Probability of being crew:
```{r}
p_crew <- titanic %>% 
  summarise(p = sum(crew == TRUE)/n())
p_crew
```
Probability of survival GIVEN being crew:

```{r}
p_survival_crew/p_crew
```

The probability is 0.237 which is the same as the first calculation.

### Calculate the probability that passengers were more likely to have survived than crew

prob( survived ∣ passenger )

```{r}
p_survived_given_passenger <- titanic %>% 
  filter(crew == FALSE) %>% 
  summarise(p = sum(survived == "yes")/n())

p_survived_given_passenger
```

The probability of surviving GIVEN being a passenger is 0.379. This is a higher probability than the probability of surviving given being crew (0.237).

Therefore the journalist was incorrect and if I have calculated my probability correctly, passengers are more likely to have survived than crew.

 - can also calculate this using the contingency table from the janitor package:
 
```{r}
library(janitor)
titanic %>%
  tabyl(crew, survived) %>%
  adorn_totals(where = c("row", "col")) %>%
  adorn_percentages(denominator = "all") %>%
  adorn_rounding(digits = 4) %>%
  adorn_title()
```
 

#2 MVP and Extension

#### 1. If I roll a six-sided dice once, what is the probability of getting a number less than 5? <br>

There are 6 outcomes for a dice roll, probability of each outcome is 1/6.

Probability of getting a number less than 5 is 4 out of those 6 possible outcomes. Therefore there is a 2/3 or 0.666 probability of getting a number less than 5.

in code:

```{r}
one_die <- rolldie(1)

one_die %>% 
  filter(X1 < 5) %>% 
  summarise(p = n()/nrow(one_die))
```


#### 2.How many outcomes are there in the sample space of an experiment in which we toss three dice?

Let's use the prob package for this.

```{r}
rolldie(3) %>% 
  nrow()
```
An alternative:
```{r}
rolldie(3) %>% 
  summarise(size_of_s = n())
```
 
216 different outcomes for a sample space in which we are throwing 3 dice at once.

#### 3. In an experiment where we roll four dice:

- how many outcomes are in the event where the total of the four dice equals 5, 6 or 7)?
- what is the probability of this event?
- what assumptions do we apply in answering part ii?

```{r}
four_dice <- rolldie(4)

#add a total column
four_dice <- four_dice %>% 
  mutate(total = rowSums(.))

#number of outcomes where total = 5,6 or 7
num_outcome_5_6_7 <- four_dice %>% 
  filter(total >= 5,
         total <= 7) %>% 
  summarise(num_outcome_5_6_7 = n())
```
There are 34 outcomes where the total of four dice rolls are equal to 5, 6 or 7.

To answer the probability of the event:
```{r}
p_5_6_7 <- num_outcome_5_6_7/nrow(four_dice)
p_5_6_7
```

The probability of the event is 0.0262. I am assuming that the dice rolls are independent events as in the outcome for one dice roll does not depend on the outcome of the others. I am assuming that all outcomes are equally likely.


#### 4. Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

Hint This is an arrangement where order matters. Of all the possible arrangements, how many are correct?

*Answer*
This is a permutation problem.
I am assuming no replacements - there are only 5 people, once one person has sat down that is it.

All possible arrangements of people in the chairs is 5! = 120 (assuming repeats). However, there is only one arrangement where the order will be correct (excluding repeats, can only be a choice of 1 for the first, 1 for the second etc). 

Therefore there is a 1/120 (0.0083333) probability that the people will be in the correct order.

#### 5. I roll six dice: what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?

```{r}
s_6 <- rolldie(6)

s_6 %>% 
  filter(isin(s_6, c(6,5,4,3,2,1), ordered = TRUE)) %>% 
  summarise(p = n()/nrow(s_6))
```


#### what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?
####[Hint have a look at the isin() function in the prob package. For example, to filter all rows containing 6, 1, 2 in that order, you could use isin(., c(6, 1, 2), ordered = TRUE)]

```{r}
s_6 <- rolldie(6)

s_6 %>% 
  filter(isin(s_6, c(6,5,4,3,2,1), ordered = FALSE)) %>% 
  summarise(p = n()/nrow(s_6))
```

