---
title: "Homework - CIs"
output: html_notebook
---
# MVP

### Task 1

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
```

```{r}
ames <- read_csv("data/ames.csv") %>% 
  clean_names()

head(ames)
```

### Task 2

```{r}
ames %>% 
  ggplot() +
  aes(x = lot_area) +
  geom_histogram(col = "white", bins = 30)
```

Looking at the histogram for `lot_area` the data is not roughly normally distributed. There appears to be a right skew in the data. It is tricky to see on the dataset but let's load the top 10 lot_areas to have a look.

Could have used a box plot to check this! Doh.

```{r}
ames %>% 
  select(lot_area) %>% 
  slice_max(lot_area, n = 10)
```
As confirmed from the histogram, there are 4 values which are much larger than the rest of the data set.

### Task 3

```{r}
#load library
library(infer)

#generate bootstrap sampling distribution
infer_lot_area <- ames %>% 
    specify(response = lot_area) %>% 
    generate(reps = 5000, type = "bootstrap") %>% 
    calculate(stat = "mean")

#visualise
infer_lot_area %>% 
visualise() +
  labs(x = "mean(lot_area)")
```
The bootstrap sampling distribution for the `mean(lot_area)` follows a normal distribution.

### Task 4

```{r}
#get the 95% CI
ci_95 <- infer_lot_area %>% 
  get_ci(level = 0.95, type = "percentile")

#add this to the plot
infer_lot_area %>% 
visualise() +
  labs(x = "mean(lot_area)") +
  shade_ci(endpoints = ci_95)
```
The bootstrap distribution of `mean(lot_area)` shows a normal distribution. The 95% confidence interval for the `mean(lot_area)` is shown as the green shaded area.

### Task 5

```{r}
#get the 99% CI
ci_99 <- infer_lot_area %>% 
  get_ci(level = 0.99, type = "percentile")

#add this to the plot
infer_lot_area %>% 
visualise() +
  labs(x = "mean(lot_area)") +
  shade_ci(endpoints = ci_99)
```
The 99% CI is shown in green above. This is broader than the 95% CI. This makes sense as it is now representing a larger interval from your data (99% as opposed to 95%) and so should be broader and encompass more data.

## Task 6

```{r}
#take the mean from the bootstrapped data
infer_mean_lot_area <- infer_lot_area %>% 
  summarise(mean = mean(stat))
infer_mean_lot_area

#just for sanity check the mean from the data
sample_mean_lot_area <- ames %>% 
  summarise(mean = mean(lot_area))
sample_mean_lot_area
```

The point estimate of the mean(lot_area) is 10149.31.

# Extension

### Task 1
Calculate a point estimate and 95% CI for the proportion of houses in the data built before 1920. Does the number of reps you use matter? [Investigate reps from 200 up to 50000, memory of your laptop permitting].

```{r}
#### WARNING - this code may run for a while! 

#create column for filtering proportion
before_1920_ames <- ames %>% 
  mutate(
    before_1920 = if_else(year_built <1920, TRUE, FALSE)
  ) %>% 
  select(year_built, before_1920)

ci_reps <- c(200, 500, 1000, 2500, 5000, 10000, 15000, 20000, 25000, 50000)
result_table <- tibble()

for (i in ci_reps){

#log value of i in console
cat(i, sep="\n")

#generate bootstrap sampling distribution with i reps from ci_reps
infer_before_1920 <- before_1920_ames %>% 
    specify(response = before_1920, success = as.character(TRUE)) %>% 
    generate(reps = i, type = "bootstrap") %>% 
    calculate(stat = "prop")

#create CI
ci_95_before_1920 <- infer_before_1920 %>% 
  get_ci(level = 0.95, type = "percentile")

#prepare result for binding to the table
result_row <- c(
  sample_num = i,
  lower = ci_95_before_1920 %>% 
    pull(lower_ci),
  upper = ci_95_before_1920 %>% 
    pull(upper_ci)
)

#add the result line to the table
result_table <- result_table %>% 
  bind_rows(result_row)

}

result_table
```

The more reps you use the smaller the confidence interval and the closer you get to approximating the population mean. I would say the improvement in values is so small that using a bootstrap of 5000 would be suitable for most applications.

Notes: forgot to plot distribution and CI! It is assumed that 10000 bootstrap reps are more reliable.